---
title: "Spaceship Titanic"
author: "Yafee Ishraq"
date: "2022-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install packages

```{r,eval=FALSE}

install.packages("data.table")
install.packages("tidyverse")
install.packages("flextable")
install.packages("ggcorrplot")


```

# Import packages

```{r,eval=TRUE}
library("data.table")
library("tidyverse")
library("flextable")
library("ggcorrplot")
```

# Unzip file
```{r,eval=FALSE}

unzip("spaceship-titanic.zip")
                    
```

# Read the data
```{r,eval=TRUE}

data_train <- fread("train.csv")
data_test  <- fread("test.csv")
sample_submission <- fread("sample_submission.csv")

```


# Data wrangling and exploratory analysis

## Missing value analysis

### Transform quotes to NA
```{r,eval=TRUE}

# Transform quotes to NA value
data_train[,colnames(data_train):=lapply(.SD,function(x)ifelse(is.na(x)|x=="",NA,x)),.SDcols = colnames(data_train)]


```


## Missing value analysis
```{r,eval=TRUE}

# Classify class for each column
col_class <- sapply(data_train,class)

col_class <- data.table(Column = names(col_class),
                        Class = col_class)

# Checking if the value is NA or not on each column
na_values <- data_train[,lapply(.SD,
                    function(x)is.na(x)),
           .SDcols = colnames(data_train)]

# Number of NA values per column
na_values_col <- na_values[,lapply(.SD,function(x)sum(x)),
                           .SDcols = colnames(na_values)]

# Pivot longer
na_values_col <- melt(na_values_col,measure.vars = colnames(na_values_col),
variable.name = "Column",value.name = "Number of missing values")

# Join column class
na_values_col <- merge(na_values_col,
                       col_class,
                       by.x = "Column",
                       by.y = "Column",
                       all.x = TRUE)

# Sort
setorder(na_values_col,-`Number of missing values`)

# Print table
na_values_col %>% 
  flextable() %>%
  set_caption("Number of missing values per column")
  
# Compute number of missing values per row
na_values[,sum_na:=rowSums(.SD),.SDcols = colnames(na_values)] 

na_values_row <- data.table(`Number of complete rows` = na_values[sum_na==0,.N],
`Number of uncomplete rows` = na_values[sum_na>0,.N])
       
# Print table                         
na_values_row %>% 
  flextable() %>% 
  colformat_int(big.mark = " ")


```

## Preprossesing data

```{r,eval=TRUE}

# ------------- Training dataset ---------------------------------- 
data_train[,VIP:=ifelse(VIP==TRUE,"VIP service","No VIP service")]
data_train[,CryoSleep:=ifelse(CryoSleep==TRUE,"Cryogenic sleep",
                              "No cryogenic sleep")]


# Transform to factors
data_train[,HomePlanet:=as.factor(HomePlanet)]
data_train[,CryoSleep:=as.factor(CryoSleep)]
data_train[,Destination :=as.factor(Destination )]
data_train[,VIP:=as.factor(VIP)]


# ------------- Test dataset ---------------------------------- 

data_test[,VIP:=ifelse(VIP==TRUE,"VIP service","No VIP service")]
data_test[,CryoSleep:=ifelse(CryoSleep==TRUE,"Cryogenic sleep",
                              "No cryogenic sleep")]


# Transform to factors
data_test[,HomePlanet:=as.factor(HomePlanet)]
data_test[,CryoSleep:=as.factor(CryoSleep)]
data_test[,Destination :=as.factor(Destination )]
data_test[,VIP:=as.factor(VIP)]

# Split cabin column into 3 columns
data_train[,c("Deck","Num","Side"):=tstrsplit(Cabin,"/")]

```


## Analyzing the target variable

```{r,eval=TRUE}

# How many got transported 
data_train %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>% 
  group_by(Transported) %>% 
  summarise(`Number of observations` = n()) %>% 
  ggplot(aes(x = Transported,
             y = `Number of observations`))+
  geom_bar(stat = "identity")+
   geom_text(aes(label = `Number of observations`),
             vjust = 2)+
  xlab("")+
  theme_bw()

plot_function <- function(x){
  
  
  data_train %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  group_by(get(x),Transported) %>% 
  summarise(Gordex = n()) %>% 
  ggplot(aes_string(x = "Transported",
             y = "Gordex",
             fill = x))+
  geom_bar(stat = "identity")+
  xlab("")+
  scale_y_continuous(n.breaks=6)+
  theme_bw()
  
}

# How many got transported, grouped by HomePlanet 
data_train %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  group_by(HomePlanet,Transported) %>% 
  summarise(`Number of observations` = n()) %>% 
  ggplot(aes(x = Transported,
             y = `Number of observations`,
             fill = HomePlanet))+
  geom_bar(stat = "identity")+
  xlab("")+
  scale_y_continuous(n.breaks=6)+
  theme_bw()

# How many got transported, grouped by destination 
data_train %>% 
  filter(!is.na(Destination)) %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  group_by(Destination,Transported) %>% 
  summarise(`Number of observations` = n()) %>% 
  ggplot(aes(x = Transported,
             y = `Number of observations`,
             fill = Destination))+
  geom_bar(stat = "identity")+
  xlab("")+
  scale_y_continuous(n.breaks=6)+
  theme_bw()

# How many got transported, grouped by destination 
data_train %>% 
  filter(!is.na(VIP)) %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  group_by(VIP,Transported) %>% 
  summarise(`Number of observations` = n()) %>% 
  ggplot(aes(x = Transported,
             y = `Number of observations`,
             fill = VIP))+
  geom_bar(stat = "identity")+
  xlab("")+
  scale_y_continuous(n.breaks=6)+
  theme_bw()

# How many got transported, grouped by cryosleep
data_train %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  group_by(CryoSleep,Transported) %>% 
  summarise(`Number of observations` = n()) %>% 
  ggplot(aes(x = Transported,
             y = `Number of observations`,
             fill = CryoSleep))+
  geom_bar(stat = "identity")+
  xlab("")+
  scale_y_continuous(n.breaks=6)+
  theme_bw()

# How many got transported, grouped by cryosleep
data_train %>% 
  filter(!is.na(Deck)) %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  group_by(Deck,Transported) %>% 
  summarise(`Number of observations` = n()) %>% 
  ggplot(aes(x = Deck,
             y = `Number of observations`,
             fill = Transported))+
  geom_bar(stat = "identity")+
  xlab("")+
  scale_y_continuous(n.breaks=7)+
  theme_bw()

# Density on age
data_train %>% 
  mutate(Transported=if_else(Transported==TRUE,"Transported","Not transported")) %>%
  ggplot(aes(x = Age,fill = Transported))+
  # facet_wrap(.~Transported)+
  geom_histogram()+
  ylab("Frequency")+
  theme_bw()


```

## Fill missing values

For numeric variables, use median.
For categorical variables, use mode.

```{r,eval=TRUE}

# Obtain numerical columns
numeric_col <- na_values_col[`Number of missing values`!=0 & Class=="numeric"]$Column

median_numeric_col<- data_train[,lapply(.SD,function(x)mean(x,na.rm=TRUE)),.SDcols = numeric_col]

data_train[,(na_values_col$Column):=lapply(.SD,function(x)ifelse(is.na(x),median(x,na.rm = TRUE),x)),.SDcols = (na_values_col$Column)]

```


## Correlation analysis
```{r,eval=TRUE}

# Classify class for each column
col_class <- sapply(data_train,class)

# Extract numerical columns
col_class_numeric <- col_class[col_class=="numeric"]

cor_matrix <- cor(data_train[,.SD,.SDcols = names(col_class_numeric)])

# Vizualise correlation matrix. Lower triangular
ggcorrplot(cor_matrix,type = "lower")

```

# Fit a classification model
```{r,eval=TRUE}

# Transform to 1 and 0.
data_train[,Transported:=ifelse(Transported==TRUE,1,0)]

logistic_reg <- glm(formula = Transported ~ HomePlanet + CryoSleep+Destination + Age, family = "binomial", 
data = data_train)

summary(logistic_reg)

# Prediction on the test dataset
test_predicton <- predict(logistic_reg,data_test[,.(HomePlanet,CryoSleep,
                                                    Destination,Age)])

# Add the prediction to the submission file. Using a cutoff on 0.5



```

